---
title: "R Notebook"
output: html_notebook
---

```{r, echo = TRUE, message = FALSE, warning = FALSE}  
#-----load required packages-----  
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,
               here,
               readxl,
               janitor) # for lemmatisation and stemming    
```

# Dataset background  

## 2020-05_DESE_jobseekers.xlsx  

JSCI data contains those in jobactive, new employment services trial, online employment services and online employment services trial at 31 May 2020 who have a valid JSCI and address as at 30 July 2020.

Sheets of interest:
+ 31 May JSCI Demographics by ADII region (Australian Digital Inclusion Index)   
+ What have you MOSTLY been doing for the last two years?  
+ In your most recent job, how many hours did you mostly work per week?  
+ JSCI Demographics by LGA  

```{r}
#-----load May 2020 job seeker data-----
jsci_all_raw_data <- read_excel(here("01_raw_data", "2020-05_DESE_jobseekers.xlsx"),
                            sheet = "JSCI demograhpics - ADII region", 
                            skip = 2) 

jsci_work_raw_data <- read_excel(here("01_raw_data", "2020-05_DESE_jobseekers.xlsx"),
                            sheet = "Q2 mostly doing - ADII region", 
                            skip = 2)

jsci_hours_data <- read_excel(here("01_raw_data", "2020-05_DESE_jobseekers.xlsx"),
                            sheet = "Q3 hours worked - ADII region", 
                            skip = 2)

jsci_lga_raw_data <- read_excel(here("01_raw_data", "2020-05_DESE_jobseekers.xlsx"),
                            sheet = "JSCI demographics - LGA", 
                            skip = 2)
```


## 2020-05_DESE_jobseekers.xlsx    
```{r}

```

